---
title: "Project Checkpoint 2"
author: "Aryan Ajith Nair"
format: html
---

## Checkpoint 2

### **Loading in the data sets**
```{r}

library(tidyverse)

cases_month <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_month.csv')
cases_year <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_year.csv')


```

### **Research question:** Do measles cases show a meaningful seasonal pattern (Do months consistently produce more cases globally?)? 

### **Creating Custom Function for Summary + Modifying Columns**

```{r}
monthly_profile <- function(x) {
  tibble(
    n = sum(!is.na(x)),
    median_cases = round(median(x, na.rm = TRUE), 1),
    mean_cases = round(mean(x, na.rm = TRUE), 1),
    total_cases = sum(x, na.rm = TRUE)
  )
}

seasonal_summary <- cases_month |>
  mutate(
    month_name = factor(month.abb[month], levels = month.abb),
    season = case_when(
      month %in% 3:5 ~ "Spring",
      month %in% 6:8 ~ "Summer",
      month %in% 9:11 ~ "Autumn",
      .default = "Winter"
    )
  ) |>
  group_by(month_name, season) |>
  reframe(monthly_profile(measles_total)) |>
  mutate(pct_of_total = round(total_cases / sum(total_cases) * 100, 1))

print(seasonal_summary)
```


### **Conclusion**

According to my summary table, measles cases peak during the spring months of March, April, and May. These 3 months account for a disproportionately large share of annual global cases. This suggests that school-term congregation after winter breaks drives transmission more than any other seasonal factor.
